---
title: "DDSAnalytics Employee Attrition Analysis"
author: "Blake Holmes" 
date: August 18, 2019
output:
  beamer_presentation:
    theme: "Berlin"
---

```{r setup_ppt, include=FALSE}
library(knitr)
library(kableExtra)
knitr::opts_chunk$set(echo = FALSE)

#pull in other Rmarkdown files to get data 
ksource <- function(x, ...) {
  library(knitr)
  source(purl(x, output = tempfile(),quiet=TRUE), ...)
}

```

```{r get_analysis, include=FALSE, echo=FALSE, cache=FALSE}
#IMPORT ANALYSIS
ksource("./Holmes_Blake_CaseStudy2_Analysis.Rmd")
```

## Agenda

### An Analysis of Employee Attrition

- Understand Employee Attrition
- Predict future attrition & salaries
- Next Steps
- link to Youtube Video [Youtube Video](https://youtu.be/6x3NGn5NUvs)

## The Data Set

```{r explanatory data, echo=FALSE}

total_rows<-nrow(att_pruned)

total_attrition_perc<-100*round(sum(att_pruned$Attrition==1)/nrow(att_pruned),2)
```
There were a total of `r total_rows` employees observed in the analysis, of which `r total_attrition_perc`% were lost to attrition. 

<br />
```{r variable_type_table, echo=FALSE}

variable_type_table %>%
  kable() %>% 
  kable_styling("striped") %>%
  row_spec(0, bold = "T")
```

## Attrition Classifier Tool
```{r, echo=FALSE}
acc_table<-ensembleClassifyResults$Accuracies %>% 
    data.frame() 

acc_table %>% 
  rename("Accuracy"=".") %>% 
  mutate(Model=str_to_title(rownames(acc_table))) %>%
  select(Model,everything()) %>% 
  arrange(desc(Accuracy)) %>% 
  mutate("Rank"=row_number()) %>%
  select("Rank","Accuracy", everything()) %>%
  kable()%>% 
  kable_styling("striped") %>%
  row_spec(0, bold = "T")

```
`r ensembleClassifyResults$WINNER_DESCRIPTION` 

## Attrition Top 6 Important Features
```{r, echo=FALSE}
head(ensembleClassifyResults$WINNER_IMPORTANCE, addrownums = TRUE) %>%
  mutate("Rank"=row_number()) %>%
  select("Rank", everything()) %>% 
  kable()%>% 
  kable_styling("striped") %>%
  row_spec(0, bold = "T")
```

## Attrition Bottom 6 Important Features
```{r, echo=FALSE}
feature_count<-nrow(ensembleClassifyResults$WINNER_IMPORTANCE)

tail(ensembleClassifyResults$WINNER_IMPORTANCE) %>% 
  as_tibble()%>%
  mutate(Rank=(feature_count-5):feature_count) %>% 
  select("Rank", everything()) %>% 
  kable() %>% 
  kable_styling("striped") %>%
  row_spec(0, bold = "T")
```

## Salary Projections Tool

```{r, echo=FALSE}
rmse_table<-ensembleRegressionResults$RMSE_results %>% 
    data.frame() 

rmse_table %>% 
  rename("Root Mean-Squared Error"=".") %>% 
  mutate(Model=rownames(rmse_table)) %>%
  select(Model,everything()) %>% 
  arrange(`Root Mean-Squared Error`) %>%
  mutate("Rank"=row_number()) %>%
  select("Rank", everything(),Model) %>% 
  mutate(`Root Mean-Squared Error`=paste0("$",round(`Root Mean-Squared Error`,0))) %>%
  kable()%>% 
  kable_styling("striped") %>%
  row_spec(0, bold = "T")
```
`r ensembleRegressionResults$WINNER_DESCRIPTION` 

## Salary Top 6 Important Features 
```{r, echo=FALSE}
head(ensembleRegressionResults$WINNER_IMPORTANCE) %>% 
  mutate("Rank"=row_number()) %>%
  select("Rank", everything()) %>% 
  kable()%>% 
  kable_styling("striped") %>%
  row_spec(0, bold = "T")
```

## Salary Bottom 6 Important Features 
```{r, echo=FALSE}
tail(ensembleRegressionResults$WINNER_IMPORTANCE) %>%
  mutate(Rank=(feature_count-5):feature_count) %>% 
  select("Rank", everything()) %>% 
  kable()%>% 
  kable_styling("striped") %>%
  row_spec(0, bold = "T")
```

## Employee Salary Histogram by Attrition

```{r, echo=FALSE,fig.height = 8, fig.align='center'}

ggplot(att_pruned, aes(x=MonthlyIncome, fill = Attrition)) + 
   geom_histogram(alpha = 0.5, aes(), position = 'identity',bins=30) + 
    scale_fill_discrete(name = "Attrition", labels = c("No", "Yes")) + 
    theme(text =element_text(size = 30), plot.margin=unit(c(0,0,6,0),"cm")) + 
    xlab("Monthly Income ($)") 

```

## Salary Compared by Attrition

- Employees retained had a mean salary of $`r TTEST_ATT_SALARY$estimate[1]`

- Employees lost due to attrition had a mean salary of $`r round(TTEST_ATT_SALARY$estimate[2])`
</br>

```{r, echo=FALSE}
salaryAttrition %>%
  group_by(Attrition) %>%
  summarise(mean(predictMinusObservered, na.rm=TRUE)) %>% 
  rename("Actual vs Predicted Income($)"=2) %>%
  mutate("Actual vs Predicted Income($)"= -1*round(`Actual vs Predicted Income($)`,2)) %>%  #*-1 to make it actuals subtract predicted
  mutate(Attrition=ifelse(Attrition==1,"Yes","No")) %>% kable()
```

## Predicted Salary versus Actual Plot

```{r, echo=FALSE}
salary_predictVSactual_attritionPLOT + 
  xlab("Predicted Monthly Income($)") + 
  ylab("Actual Monthly Income($)")+ scale_color_discrete(name = "Attrition", labels = c("No","Yes")) + 
  theme(plot.margin=unit(c(0,0,5,0),"cm")) 
```

## Attrition in Protected Classes 

```{r protected class, echo=FALSE}

protectedClassesPLOT + theme(text =element_text(size = 15), plot.margin=unit(c(0,0,6,0),"cm")) +scale_x_discrete(labels=c("0" = "No", "1" = "Yes"))

```
[Equal Employment Opportunity Commision](https://www.eeoc.gov/laws/types/)

## Attrition in Protected Classes 

```{r protected class 2, echo=FALSE}
protectedClassStats %>% rename(Female=pct_female,
                               "Over 40"=over_40) %>% 
                                mutate(Attrition=ifelse(Attrition==1,"Yes","No")) %>% 
                                mutate('Average Age'=round(avg_age,1)) %>% 
                                select(-avg_age) %>% 
                                kable()%>% 
                                kable_styling("striped") %>%
                                row_spec(0, bold = "T")
                                
```

## Future Projections

`r FUTURE_SALARIES_DESCRIPTION` 

`r FUTURE_ATT_DESCRIPTION`

## Next Steps

1. Pilot attrition intervention program on "at-risk" employees.
2. Gather more subjective or perception-based data.
3. Further analysis and Data Modeling.